A Sample UBC Thesis

### 2014-04-14 thesis-template/ubcsample.tex
thunk

%% LyX 1.6.7 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english,msc]{ubcthesis}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{longtable}
\usepackage{url}
\usepackage{graphicx}
\usepackage{esint}
\usepackage[numbers]{natbib}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
%%
%% This is file `ubcsample.tex',
%% generated with the docstrip utility.
%
% The original source files were:
%
% ubcthesis.dtx  (with options: `ubcsampletex')
%% 
%% This file was generated from the ubcthesis package.
%% --------------------------------------------------------------
%% 
%% Copyright (C) 2001
%% Michael McNeil Forbes
%% mforbes@alum.mit.edu
%% 
%% This file may be distributed and/or modified under the
%% conditions of the LaTeX Project Public License, either version 1.2
%% of this license or (at your option) any later version.
%% The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% This program is distributed in the hope that it will be useful,
%% but WITHOUT ANY WARRANTY; without even the implied warranty of
%% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
%% LaTeX Project Public License for more details.
%% 
%% This program consists of the files ubcthesis.dtx, ubcthesis.ins, and
%% the sample figures fig.eps and fig.fig.
%% 
%% This file may be modified and used as a base for your thesis without
%% including the licence agreement as long as the content (i.e. textual
%% body) of the file is completely rewritten. You must, however, change
%% the name of the file.
%% 
%% This file may only be distributed together with a copy of this
%% program. You may, however, distribute this program without generated
%% files such as this one.
%%

% This Sample thesis requires \LaTeX2e
\NeedsTeXFormat{LaTeX2e}[1995/12/01]
\ProvidesFile{ubcsample.tex}[2010/06/13 v1.66 ^^J
 University of British Columbia Sample Thesis]
% This is the \documentclass[]{} command.  The manditory argument
% specifies the "flavour" of thesis (ubcthesis for UBC).  The
% optional arguments (in []) specify options that affect how the
% thesis is displayed.  Please see the ubcthesis documentation for
% details about the options.

%
% To compile this sample thesis, issue the following commands:
% latex ubcsample
% bibtex ubcsample
% latex ubcsample
% latex ubcsample
% latex ubcsample
%
% To view use xdvi (on unix systems):
% xdvi ubcsample.dvi
%
% To make a postscript file, use dvips:
% dvips -o ubcsample.ps ubcsample.dvi
%
% To view the postscript file, use ghostview or gv (on unix systems):
% gv ubcsample.ps
%
%************************************************
% Optional packages.
%
% The use of these packages is optional, but they provide various
% tools for more flexible formating.  The sample thesis uses these,
% but if you remove the example code, you should be able to exclude
% these packages.  Only standard packages have been described here;
% they should be installed with any complete LaTeX instalation, but
% if not, you can find them at the Comprehensive TeX Archive Network
% (CTAN): http://www.ctan.org/
%

@@

%******** graphics and graphicx *****************
% This allows you to include encapsulated postscript files.  If you
% don't have this, comment the \includegraphics{} line following the
% comment "%includegraphics" later in this file.

@@

%******** natbib ********************************
% This is a very nice package for bibliographies.  It includes options
% for sorting and compressing bibliographic entries.

@@

% These commands are required.
\title{A Sample UBC Thesis}
\subtitle{With a Subtitle}
\author{Michael M$^{\rm c}$Neil Forbes}
\copyrightyear{2000}
\submitdate{\today} % The "\ " is required after
                                      % \monthname to prevent the
                                      % command from eating the space.
\program{Physics}

@@

% One might want to override the format of the section and chapter
% numbers.  This shows you how to do it.  Note that the current
% format is acceptable for submission to the FoGS: If you wish to modify
% these, you should check with the FoGS explicity. prior to making
% the modifications.
\renewcommand{\thepart}{\Roman{part}}
\renewcommand{\thechapter}{\arabic{chapter}}
\renewcommand{\thesection}{\thechapter.\arabic{section}}
\renewcommand{\thesubsection}{\thesection.\arabic{subsection}}
\renewcommand{\thesubsubsection}{\thesubsection.\arabic{subsubsection}}
\renewcommand{\theparagraph}{\thesubsubsection.\arabic{paragraph}}
\renewcommand{\thesubparagraph}{\theparagraph.\arabic{subparagraph}}

@@

% Here is the start of the document.

\makeatother

\usepackage{babel}

\begin{document}
% This starts numbering in Roman numerals as required for the thesis
% style.
\frontmatter % Mandatory

% The order of the following components should be preserved.  The order
% listed here is the order currently required by FoGS.
% Mandatory
% Mandatory -  maximum 350 words

\begin{abstract}
The \texttt{genthesis.cls} \LaTeX{} class file and accompanying documents,
such as this sample thesis, are distributed in the hope that it will
be useful but without any warranty (without even the implied warranty
of fitness for a particular purpose). For a description of this file's
purpose, and instructions on its use, see below.

@@

Michael Forbes. mforbes@physics.ubc.ca 
\end{abstract}
\tableofcontents{}% Mandatory
\listoftables
% Mandatory if thesis has tables
\listoffigures
% Mandatory if thesis has figures
\listof{Program}{List of Programs} % Optional
% Any other lists should come here, i.e.
% Abbreviation schemes, definitions, lists of formulae, list of
% schemes, glossary, list of symbols etc.

\chapter{Acknowledgements}

% Optional
This is the place to thank professional colleagues and people who
have given you the most help during the course of your graduate work.

\chapter{Dedication}

% Optional
The dedication is usually quite short, and is a personal rather than
an academic recognition. The \emph{Dedication} does not have to be
titled, but it must appear in the table of contents. If you want to
skip the chapter title but still enter it into the Table of Contents,
use this command \verb|\chapter[Dedication]{}|.

\chapter{Statement of Co-Authorship}

% Manditory if thesis co-written
If any part of your thesis was co-written, you must include a Co-Authorship
statement which gives details of your contribution to the following
areas:
\begin{itemize}
\item identification and design of research program 
\item performing the research 
\item data analyses 
\item manuscript preparation 
\end{itemize}
Note that this section is the last of the preliminary pages (with
lowercase Roman numeral page numbers). It must be placed \emph{before}
the \verb|\mainmatter| command. After that, Arabic numbered pages
will begin.

@@

% Chapters are the next main unit.

\chapter{This is a Chapter}

% Sections are a sub-unit

\section{A Section}

Here is a section with some text. Equations look like this $y=x$.%
\footnote{Here is a footnote.%
}

@@

% Subsections follow

\subsection{This is a Subsection}

Here is an example of a citation: \citep{Forbes:2006ba}. The actual
form of the citation is governed by the bibliographystyle. These citations
are maintained in a BIB\TeX{} file \texttt{sample.bib}. You could
type these directly into the file. For an example of the format to
use look at the file \texttt{ubcsample.bbl} after you compile this
file.%
\footnote{Here is another footnote.%
}

@@

%
\begin{table}[t]
 %optional [t, b or h];
 \begin{tabular}{|r||r@{.}l}
\hline 
Phoenix  & \$960 & 35\tabularnewline
\hline 
Calgary  & \$250 & 00\tabularnewline
\hline
\end{tabular}\caption[Here is the caption for this wonderful table\ldots{}]{ \label{tab:Table1} Here is the caption for this wonderful table.
It has not been centered and the positioning has been specified to
be at the top of the page. Thus it appears above the babble rather
than below where it is defined in the source file.}

\end{table}

% Force a new page: without this, the quote would appear on the
% previous page.
\newpage{}

\section{Quote}

% It is centered

\begin{quote}
\begin{center}
This is a small poem,\\
 a little poem, a Haiku,\\
 to show you how to.\\
 ---Michael M$^{{\rm c}}$Neil Forbes. 
\par\end{center}
\end{quote}
This small poem shows several features: 
\begin{itemize}
\item The use of the \verb|quote| and \verb|center| environments. 
\item The \verb|\newpage| command has been used to force a page break.
(Sections do not usually start on a new page.) 
\item The pagestyle has been set to suppress the headers using the command
\verb|\thispagestyle{plain}|. Note that using \verb|\pagestyle{plain}|
would have affected all of the subsequent pages. 
\end{itemize}

\section{Programs}

Here we give an example of a new float as defined using the \texttt{float}
package. In the preamble we have use the commands \begin{verbatim}
\floatstyle{ruled} \newfloat{Program}{htbp}{lop}{[}chapter{]}
\end{verbatim} This creates a {}``Program'' environment that may
be use for program fragments. A sample \texttt{python} program is
show as Program~\ref{prog:fib}. (Note that Python places a fairly
restrictive limit on recursion so trying to call this with a large
$n$ before building up the cache is likely to fail unless you increase
the recursion depth.) \begin{Program} \caption{\label{prog:fib} Python program that computes the $n^{{\rm th}}$
Fibonacci number using memoization.}

\begin{verbatim} def fib(n,cache={}): if n < 2: return 1 if n in
cache: return cache{[}n{]} else: result = fib(n-1)+fib(n-2) cache{[}n{]}
= result return result \end{verbatim} \end{Program} Instead of using
a \texttt{verbatim} environment for your program chunks, you might
like to \texttt{include} them within an \texttt{alltt} envrironment
by including the \verb|\usepackage{alltt}| package (see page 187
of the \LaTeX{} book). Another useful package is the \verb|\usepackage{listings}|
which can pretty-print many different types of source code.

% Force a new page
\newpage{}

% Here we provide a short optional argument to \chapter[]{}.  This
% optional argument will appear in the table of contents.  For long
% titles, one should use this to give a single-line entry to the
% table of contents.

\chapter[Another Chapter\ldots{}]{Another Chapter with a Very Long Chapter-name that will Probably
Cause Problems}

@@

\section{Another Section}

Another bunch of text to demonstrate what this file does. You might
want a list for example:%
\footnote{Here is a footnote in a different chapter. Footnotes should come after
punctuation.%
} 
\begin{itemize}
\item An item in a list. 
\item Another item in a list. 
\end{itemize}

\section*{An Unnumbered Section That is Not Included in the Table of Contents}

Here is an example of a figure environment. %
\begin{figure}[ht]
 %psfrag: comment the following line if not using the psfrag package

\centering{}\psfrag{pie makes me happy!}{$\pi$ makes me happy!}
%includegraphics: comment the following if not using the graphicx package
 \includegraphics[width=0.4\textwidth]{fig} \caption[Happy Face: figure example.]{\label{fig:happy} This is a figure of a happy face with a \texttt{psfrag}
replacement. The original figure (drawn in xfig and exported to a
.eps file) has the text {}``pie makes me happy!''. The \texttt{psfrag}
package replaces this with {}``$\pi$ makes me happy!''. Note that
we have used the optional argument for the caption command so that
only a short version of this caption occurs in the list of figures.}

\end{figure}

\afterpage{\clearpage{}} Perhaps I should say that the example
of a figure can be seen in Figure~\ref{fig:happy}. Figure placement
can be tricky with \LaTeX{}\ because figures and tables are treated
as {}``floats'': text can flow around them, but if there is not
enough space, they will appear later. To prevent figures from going
too far, the \verb|\afterpage{\clearpage}| command can be used. This
makes sure that the figure appears on the following page.

@@

Figures can make a document more enjoyable as demonstrated by Figure~\ref{fig:happy}.

\section{Tables}

We have already included one table:~\ref{tab:Table1}. Another table
is plopped right here. %
\begin{table}[ht]

\centering{}\begin{tabular}{|l||l|l||l|l|}
\hline 
 & \multicolumn{2}{l|}{Singular} & \multicolumn{2}{l|}{Plural}\tabularnewline
\hline 
 & English & \textbf{Gaeilge} & English & \textbf{Gaeilge}\tabularnewline
\hline
\hline 
1st Person & at me & \textbf{agam} & at us & \textbf{againn}\tabularnewline
2nd Person & at you & \textbf{agat} & at you & \textbf{agaibh}\tabularnewline
3rd Person & at him & \textbf{aige} & at them & \textbf{acu}\tabularnewline
 & at her & \textbf{aici} &  & \tabularnewline
\hline
\end{tabular}\caption{ \label{tab:Table2} Another table.}

\end{table}

Well, actually, as with Figure~\ref{fig:happy}, the table does not
necessarily appear right {}``here'' because tables are also {}``floats''.
\LaTeX{} puts them where it can. Becuase of this, one should refer
to floats by their labels rather than by their location. This example
is demonstrated by Table~\ref{tab:Table2}. This one is pretty close,
however.

Another useful package is \verb|\usepackage{longtable}| which provides
the \texttt{longtable} environment. This is nice because it allows
tables to span multiple pages. Table~\ref{tab:longtable} has been
formatted this way.

\begin{center}
\begin{longtable}{|l|l|l|}
\caption{\label{tab:longtable}Feasible triples for highly variable Grid}
\tabularnewline
\caption{\label{tab:longtable}Feasible triples for highly variable Grid}
\tabularnewline
\hline 
\multicolumn{1}{|c|}{\textbf{Time (s)}} & \multicolumn{1}{c|}{\textbf{Triple chosen}} & \multicolumn{1}{c|}{\textbf{Other feasible triples}}\tabularnewline
\endfirsthead
\hline 
\multicolumn{3}{c}{}\tabularnewline
\hline 
\multicolumn{1}{|c|}{\textbf{Time (s)}} & \multicolumn{1}{c|}{\textbf{Triple chosen}} & \multicolumn{1}{c|}{\textbf{Other feasible triples}}\tabularnewline
\hline
\endhead
\hline 
\multicolumn{3}{|r|}{{Continued on next page}}\tabularnewline
\hline
\endfoot
\hline 
0  & (1, 11, 13725)  & (1, 12, 10980), (1, 13, 8235), (2, 2, 0), (3, 1, 0) \tabularnewline
274  & (1, 12, 10980)  & (1, 13, 8235), (2, 2, 0), (2, 3, 0), (3, 1, 0) \tabularnewline
5490  & (1, 12, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
8235  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
10980  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
13725  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
16470  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
19215  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
21960  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
24705  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
27450  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
30195  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
32940  & (1, 13, 16470)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
35685  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
38430  & (1, 13, 10980)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
41175  & (1, 12, 13725)  & (1, 13, 10980), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
43920  & (1, 13, 10980)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
46665  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
49410  & (2, 2, 2745)  & (2, 3, 0), (3, 1, 0) \tabularnewline
52155  & (1, 12, 16470)  & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
54900  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
57645  & (1, 13, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline
60390  & (1, 12, 13725)  & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \tabularnewline

>>>>

%%
%% This is file `ubcsample.tex',
%% generated with the docstrip utility.
%
% The original source files were:
%
% ubcthesis.dtx  (with options: `ubcsampletex')
%% 
%% This file was generated from the ubcthesis package.
%% --------------------------------------------------------------
%% 
%% Copyright (C) 2001
%% Michael McNeil Forbes
%% mforbes@alum.mit.edu
%% 
%% This file may be distributed and/or modified under the
%% conditions of the LaTeX Project Public License, either version 1.2
%% of this license or (at your option) any later version.
%% The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% This program is distributed in the hope that it will be useful,
%% but WITHOUT ANY WARRANTY; without even the implied warranty of
%% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
%% LaTeX Project Public License for more details.
%% 
%% This program consists of the files ubcthesis.dtx, ubcthesis.ins, and
%% the sample figures fig.eps and fig.fig.
%% 
%% This file may be modified and used as a base for your thesis without
%% including the licence agreement as long as the content (i.e. textual
%% body) of the file is completely rewritten. You must, however, change
%% the name of the file.
%% 
%% This file may only be distributed together with a copy of this
%% program. You may, however, distribute this program without generated
%% files such as this one.
%%

% This Sample thesis requires \LaTeX2e
\NeedsTeXFormat{LaTeX2e}[1995/12/01]
\ProvidesFile{ubcsample.tex}[2012/04/07 v1.70 ^^J
 University of British Columbia Sample Thesis]
% This is the \documentclass[]{} command.  The manditory argument
% specifies the "flavour" of thesis (ubcthesis for UBC).  The
% optional arguments (in []) specify options that affect how the
% thesis is displayed.  Please see the ubcthesis documentation for
% details about the options.
\documentclass[msc,oneside]{ubcthesis}
%
% To compile this sample thesis, issue the following commands:
% latex ubcsample
% bibtex ubcsample
% latex ubcsample
% latex ubcsample
% latex ubcsample
%
% To view use xdvi (on unix systems):
% xdvi ubcsample.dvi
%
% To make a postscript file, use dvips:
% dvips -o ubcsample.ps ubcsample.dvi
%
% To view the postscript file, use ghostview or gv (on unix systems):
% gv ubcsample.ps
%
%************************************************
% Optional packages.
%
% The use of these packages is optional, but they provide various
% tools for more flexible formating.  The sample thesis uses these,
% but if you remove the example code, you should be able to exclude
% these packages.  Only standard packages have been described here;
% they should be installed with any complete LaTeX instalation, but
% if not, you can find them at the Comprehensive TeX Archive Network
% (CTAN): http://www.ctan.org/
%

@@

%******** graphics and graphicx *****************
% This allows you to include encapsulated postscript files.  If you
% don't have this, comment the \includegraphics{} line following the
% comment "%includegraphics" later in this file.
\usepackage{graphicx}

@@

%******** natbib ********************************
% This is a very nice package for bibliographies.  It includes options
% for sorting and compressing bibliographic entries.
\usepackage[numbers,sort&compress]{natbib}

@@

% These commands are required.
\title{A Sample UBC Thesis}
\subtitle{With a Subtitle}
\author{Michael M$^{\rm c}$Neil Forbes}
\copyrightyear{2000}
\submitdate{\monthname\ \number\year} % The "\ " is required after
                                      % \monthname to prevent the
                                      % command from eating the space.
\program{Physics}

@@

% One might want to override the format of the section and chapter
% numbers.  This shows you how to do it.  Note that the current
% format is acceptable for submission to the FoGS: If you wish to modify
% these, you should check with the FoGS explicity. prior to making
% the modifications.
\renewcommand\thepart         {\Roman{part}}
\renewcommand\thechapter      {\arabic{chapter}}
\renewcommand\thesection      {\thechapter.\arabic{section}}
\renewcommand\thesubsection   {\thesection.\arabic{subsection}}
\renewcommand\thesubsubsection{\thesubsection.\arabic{subsubsection}}
\renewcommand\theparagraph    {\thesubsubsection.\arabic{paragraph}}
\renewcommand\thesubparagraph {\theparagraph.\arabic{subparagraph}}

\setcounter{tocdepth}{2}
\setcounter{secnumdepth}{2}

@@

% Here is the start of the document.
\begin{document}

%% This starts numbering in Roman numerals as required for the thesis
%% style and is mandatory.
\frontmatter

%%% The order of the following components should be preserved.  The order
%%% listed here is the order currently required by FoGS:        \\
%%% Title (Mandatory)                                           \\
%%% Preface (Manditory if any collaborator contributions)       \\
%%% Abstract (Mandatory)                                        \\
%%% List of Contents, Tables, Figures, etc. (As appropriate)    \\
%%% Acknowledgements (Optional)                                 \\
%%% Dedication (Optional)                                       \\

\maketitle                      %% Mandatory
\begin{abstract}                %% Mandatory -  maximum 350 words
  The \texttt{genthesis.cls} \LaTeX{} class file and accompanying
  documents, such as this sample thesis, are distributed in the hope
  that it will be useful but without any warranty (without even the
  implied warranty of fitness for a particular purpose).  For a
  description of this file's purpose, and instructions on its use, see
  below.

@@

  Michael Forbes.
  mforbes@physics.ubc.ca
\end{abstract}

\chapter{Preface} % Manditory if any of the conditions are met

You must include a preface if any part of your research was partly or
wholly published in articles, was part of a collaboration, or required
the approval of UBC Research Ethics Boards.

The Preface must include the following:

\begin{itemize}
\item A statement indicating the relative contributions of all
  collaborators and co-authors of publications (if any), emphasizing
  details of your contribution, and stating the proportion of research
  and writing conducted by you.
\item A list of any publications arising from work presented in the
  dissertation, and the chapter(s) in which the work is located.
\item The name of the particular UBC Research Ethics Board, and the
  Certificate Number(s) of the Ethics Certificate(s) obtained, if
  ethics approval was required for the research.
\end{itemize}

%%% Sections and subsections etc. in the Preface should in general
%%% not be listed in the table of contents, so use the starred form
%%% of \section etc.
\section*{Examples}
Chapter~\ref{cha:apple_ref} is based on work conducted in UBC's Maple
Syrup Laboratory by Dr. A.  Apple, Professor B. Boat, and Michael
McNeil Forbes. I was responsible for tapping the trees in forests X
and Z, conducted and supervised all boiling operations, and performed
frequent quality control tests on the product.

A version of chapter~\ref{cha:apple_ref} has been
published~\cite{Apple:2010}. I conducted all the testing and wrote
most of the manuscript. The section on ``Testing Implements'' was
originally drafted by Boat, B.  Check the first pages of this
chapter to see footnotes with similar information.

Note that this preface must come before the table of contents.  Note
also that this section ``Examples'' should not be listed in the table
of contents, so we have used the starred form: \verb|\section*{Example}|.

\tableofcontents                %% Mandatory
\listoftables                   %% Mandatory if thesis has tables
\listoffigures                  %% Mandatory if thesis has figures
\listof{Program}{List of Programs} %% Optional
%% Any other lists should come here, i.e.
%% Abbreviation schemes, definitions, lists of formulae, list of
%% schemes, glossary, list of symbols etc.

\chapter{Acknowledgements}      %% Optional
This is the place to thank professional colleagues and people who have
given you the most help during the course of your graduate work.

\chapter{Dedication} %% Optional
The dedication is usually quite short, and is a personal rather than
an academic recognition.  The \emph{Dedication} does not have to be
titled, but it must appear in the table of contents.  If you want to
skip the chapter title but still enter it into the Table of Contents,
use this command \verb|\chapter[Dedication]{}|.

Note that this section is the last of the preliminary pages (with
lowercase Roman numeral page numbers).  It must be placed
\emph{before} the \verb|\mainmatter| command.  After that, Arabic
numbered pages will begin.

@@

% Chapters are the next main unit.
\chapter{This is a Chapter}

% Sections are a sub-unit
\section{A Section}
Here is a section with some text.  Equations look like this
$y=x$.\footnote{Here is a footnote.}

@@

% Subsections follow
\subsection{This is a Subsection}
Here is an example of a citation: \cite{Forbes:2006ba}.  The actual
form of the citation is governed by the bibliographystyle.  These
citations are maintained in a BIBTeX file \texttt{sample.bib}.  You
could type these directly into the file.  For an example of the format
to use look at the file \texttt{ubcsample.bbl} after you compile this
file.\footnote{Here is another footnote.}

@@

\begin{table}[t]                 % optional [t, b or h];
  \begin{tabular}{|r||r@{.}l|}
    \hline
    Phoenix & \$960&35\\
    \hline
    Calgary & \$250&00\\
    \hline
  \end{tabular}
  \caption[Here is the caption for this wonderful table\ldots]{
    \label{tab:Table1}
    Here is the caption for this wonderful table. It has not been
    centered and the positioning has been specified to be at the top
    of the page.  Thus it appears above the babble rather than below
    where it is defined in the source file.}
\end{table}

% Force a new page: without this, the quote would appear on the
% previous page.
\newpage

\section{Quote}
Here is a quote:
\begin{quote}
  % It is centered
  \begin{center}
    This is a small poem,\\
    a little poem, a Haiku,\\
    to show you how to.\\
    ---Michael M$^{\rm c}$Neil Forbes.
  \end{center}
\end{quote}

This small poem shows several features:
\begin{itemize}
\item The use of the \verb|quote| and \verb|center| environments.
\item The \verb|\newpage| command has been used to force a page
  break.  (Sections do not usually start on a new page.)
\item The pagestyle has been set to suppress the headers using the
  command \verb|\thispagestyle{plain}|.  Note that using
  \verb|\pagestyle{plain}| would have affected all of the subsequent
  pages.
\end{itemize}
\section{Programs}
Here we give an example of a new float as defined using the
\texttt{float} package.  In the preamble we have used the commands
\begin{verbatim}
\floatstyle{ruled}
\newfloat{Program}{htbp}{lop}[chapter]
\end{verbatim}
This creates a ``Program'' environment that may be used for program
fragments.  A sample \texttt{python} program is shown in
Program~\ref{prog:fib}.  (Note that Python places a fairly restrictive
limit on recursion so trying to call this with a large $n$ before
building up the cache is likely to fail unless you increase the
recursion depth.)
\begin{Program}
  \caption{\label{prog:fib} Python program that computes the $n^{\rm
      th}$ Fibonacci number using memoization.}
\begin{verbatim}
def fib(n,_cache={}):
    if n < 2:
        return 1
    if n in _cache:
        return _cache[n]
    else:
        result = fib(n-1)+fib(n-2)
        _cache[n] = result
        return result
\end{verbatim}
\end{Program}
Instead of using a \texttt{verbatim} environment for your program
chunks, you might like to \texttt{include} them within an
\texttt{alltt} envrironment by including the \verb|\usepackage{alltt}|
package (see page 187 of the \LaTeX{} book).  Another useful package
is the \verb|\usepackage{listings}| which can pretty-print many
different types of source code.

% Force a new page
\newpage

%% Here we provide a short optional argument to \chapter[]{}.  This
%% optional argument will appear in the table of contents.  For long
%% titles, one should use this to give a single-line entry to the
%% table of contents.
\chapter[Another Chapter\ldots]{Another Chapter with a Very Long
  Chapter-name that will Probably Cause Problems}
\label{cha:apple_ref}

@@

This Chapter was used as a demonstration in the Preface for how to
attribute contribution from collaborators.  If there are any such
contributions, details must be included in the Preface.  If you wish,
you may additionally use a footnote such as this.\footnote{This
  chapter is based on work conducted in UBC's Maple Syrup Laboratory
  by Dr. A. Apple, Professor B. Boat, and C. Cat.}

\section{Another Section}
Another bunch of text to demonstrate what this file does.
You might want a list for example:\footnote{Here is a footnote in a
  different chapter.  Footnotes should come after punctuation.}
\begin{itemize}
\item An item in a list.
\item Another item in a list.
\end{itemize}

\section*{An Unnumbered Section That is Not Included in the Table of
  Contents}
\begin{figure}[ht]
  \begin{center}
%% psfrag: comment the following line if not using the psfrag package
    \psfrag{pie makes me happy!}{$\pi$ makes me happy!}
%% includegraphics: comment the following if not using the graphicx package
    \includegraphics[width=0.4\textwidth]{fig.eps}
    \caption[Happy Face: figure example.]{\label{fig:happy} This is a figure of
      a happy face with a \texttt{psfrag} replacement.  The original figure
      (drawn in xfig and exported to a .eps file) has the text ``pie makes me
      happy!''.  The \texttt{psfrag} package replaces this with ``$\pi$ makes me
      happy!''.  Note: the Makefile compiles the sample using pdf\LaTeX\ which
      cannot use \texttt{psfrag} directly.  For some options that work with
      pdf\LaTeX, please see this discussion:
      \url{http://tex.stackexchange.com/questions/11839}.  For the caption, we
      have used the optional argument for the caption command so that only a
      short version of this caption occurs in the list of figures.}
  \end{center}
\end{figure}
\afterpage{\clearpage}
Here is an example of a figure environment.
Perhaps I should say that the example of a figure can be seen in
Figure~\ref{fig:happy}.  Figure placement can be tricky with \LaTeX\
because figures and tables are treated as ``floats'': text can flow
around them, but if there is not enough space, they will appear later.
To prevent figures from going too far, the
\verb|\afterpage{\clearpage}| command can be used.  This makes sure
that the figure are typeset at the end of the page (possibly appear on
their own on the following pages) and before any subsequent text.

@@

\section{Tables}
We have already included one table:~\ref{tab:Table1}.  Another table
is plopped right here.
\begin{table}[ht]
  \begin{center}
    \begin{tabular}{|l||l|l||l|l|}
      \hline
      &\multicolumn{2}{l|}{Singular}&\multicolumn{2}{l|}{Plural}\\
      \cline{2-5}
       &English&\textbf{Gaeilge}&English&\textbf{Gaeilge}\\
      \hline\hline
      1st Person&at me&\textbf{agam}&at us&\textbf{againn}\\
      2nd Person&at you&\textbf{agat}&at you&\textbf{agaibh}\\
      3rd Person&at him&\textbf{aige}&at them&\textbf{acu}\\
       &at her&\textbf{aici}& & \\
      \hline
    \end{tabular}
    \caption{
      \label{tab:Table2}
      Another table.}
  \end{center}
\end{table}
Well, actually, as with Figures, tables do not
necessarily appear right ``here'' because tables are also ``floats''.
\LaTeX{} puts them where it can.  Because of this, one should refer to
floats by their labels rather than by their location.  This example is
demonstrated by Table~\ref{tab:Table2}.  This one is pretty close,
however.  (Note: you should generally not put tables or figures in the
middle of a paragraph.  This example is for demonstration purposes
only.)

Another useful package is \verb|\usepackage{longtable}| which provides
the \texttt{longtable} environment.  This is nice because it allows
tables to span multiple pages.  Table~\ref{tab:longtable} has been
formatted this way.
\begin{center}
  \begin{longtable}{|l|l|l|}
    \caption{\label{tab:longtable}Feasible triples for
      highly variable Grid}\\

    \hline \multicolumn{1}{|c|}{\textbf{Time (s)}} &
    \multicolumn{1}{c|}{\textbf{Triple chosen}} &
    \multicolumn{1}{c|}{\textbf{Other feasible triples}} \\ \hline
    \endfirsthead

    \multicolumn{3}{c}%
    {{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
    \hline \multicolumn{1}{|c|}{\textbf{Time (s)}} &
    \multicolumn{1}{c|}{\textbf{Triple chosen}} &
    \multicolumn{1}{c|}{\textbf{Other feasible triples}} \\ \hline
    \endhead

    \hline \multicolumn{3}{|r|}{{Continued on next page}} \\ \hline
    \endfoot

    \hline \hline
    \endlastfoot

    0 & (1, 11, 13725) & (1, 12, 10980), (1, 13, 8235), (2, 2, 0), (3, 1, 0) \\
    274 & (1, 12, 10980) & (1, 13, 8235), (2, 2, 0), (2, 3, 0), (3, 1, 0) \\
    5490 & (1, 12, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
    8235 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
    10980 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
    13725 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
    16470 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
    19215 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
    21960 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
    24705 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
    27450 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
    30195 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
    32940 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
    35685 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
    38430 & (1, 13, 10980) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
    41175 & (1, 12, 13725) & (1, 13, 10980), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
    43920 & (1, 13, 10980) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
    46665 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
    49410 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
    52155 & (1, 12, 16470) & (1, 13, 13725), (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
    54900 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
    57645 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
    60390 & (1, 12, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
    63135 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
    65880 & (1, 13, 16470) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
    68625 & (2, 2, 2745) & (2, 3, 0), (3, 1, 0) \\
    71370 & (1, 13, 13725) & (2, 2, 2745), (2, 3, 0), (3, 1, 0) \\
